Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-06-03 thesis/chapters/ch_chapter4.tex
chapter 5 edits from mordecai. Still have figure edits to do

The elemental abundances of a galaxy over time are sensitive to the nuclear physics and stellar astrophysics that determines which stars make what elements and when. The stellar abundance patterns within a given galaxy also depend on the details of how those metals are released into the ISM through various forms of stellar feedback, and the hydrodynamic interactions that ultimately mix those elements into the ISM or eject them from the galaxy in galactic winds. Stellar abundance patterns in old, metal poor systems are the result of enrichment from substantially fewer sources than younger, more metal rich stellar populations. Because of this, studying abundance patterns in these environments allows us to place constraints on the yields of individual enrichment events and the nucleosynthetic sites of each element. The lowest mass dwarf galaxies in the Local Group, UFDs, offer some of the best constraints on these processes. These environments can potentially be used to trace and constrain enrichment from the first stars \citep[e.g.][]{FrebelBromm2012,Ji2015,Ritter2015,Jeon2017,Hartwig2018} and have been used to place constraints on the variety of possible astrophysical sources of r-process enrichment \citep[e.g.][]{Ji2016a,Ji2016b,Ji2018a,Ji2019,Tsujimoto2017,Duggan2018,Nagasawa2018,Ojima2018}. It is likely that chemical enrichment in more massive, metal rich systems with ongoing star formation like the Milky Way is much more uniform, averaging over the variety of complex hydrodynamics and feedback-driven turbulence interactions that may be driving much of the significant variance in stellar abundances seen at low metallicity. Stochasticity, which has been shown to be important for setting the width of stellar abundance patterns in low metallicity environments of the Milky Way halo \citep[e.g.][]{Cescutti2008,Cescutti2014}, plays an even greater role in the evolution of these low mass dwarf galaxies \citep{Applebaum2018}. Indeed the increase in scatter in the stellar abundance patterns of low mass dwarf galaxies and UFDs has been attributed to inhomogeneous mixing and stochastic effects \citep[e.g.][]{Norris2010a,Simon2015,Mashonkina2017,Suda2017}. This complicates the interpretation of abundance patterns in these galaxies, particularly for constraining individual enrichment events.

@@

In this Chapter we utilize the simulations analyzed in prior chapters to conduct a controlled set of ``mixing experiments'' whereby we restart each simulation with enrichment events placed by-hand in order to more directly investigate the evolution of metals from an individual event. While elemental yields were tracked for each star in our simulation, we lacked the necessary Lagrangian information about the metals once they were released into the ISM to be able to trace the evolution of single enrichment events. We investigate primarily how the feedback ejection energy of individual sources ($E_{\rm ej}$) and global star formation rate at the time of enrichment affects how metals are ejected from the galaxy in galactic winds and metal mixing in the ISM. In Section~\ref{ch4:sec:methods} we briefly outline our methods and discuss the setup of these mixing experiments. In Section~\ref{ch4:sec:results} we discuss the results of these experiments and the role of $E_{\rm ej}$, global SFR, radial position of the enrichment event, and local ISM density around each enrichment sources affects the evolution of its metals, and discuss and conclude in Section~\ref{ch4:sec:discussion conclusion}.

@@

\subsection{Mixing Experiment Setup}
\label{ch4:sec:experiment}
We restart our fiducial, full-physics simulation at two different times, 180 Myr, and 360 Myr, labeled as runs \runonenu~and \runtwonu~respectively. These correspond to two different times in the galaxy's SFR evolution, testing how much variance is expected in the metal mixing and ejection with the star formation rate. \runonenu~occurs during the lull in star formation following the initial SFR peak, and \runtwonu~occurs in an extended period of little to no ongoing star formation. We attempted to to evolve each simulation for 150 Myr, but due to computational constrains this was not always possible.\footnote{We began runs during the peak of star formation ($\sim 100$~Myr), but the computational cost increases substantially during this time. These runs were abandoned.}

@@

At the beginning of each restart, we place by-hand one or more enrichment events at assigned positions throughout the galaxy, with thermal injection energies ($E_{\rm ej}$), masses ($m_{\rm ej}$), and metal fractions ($Z_{\rm ej}$). We note that the injection masses, particularly the metal fractions, are somewhat arbitrary as they are likely dynamically insignificant relative to the ambient ISM mass in which they occur. The important parameter here is $E_{\rm ej}$, which we vary to sample the range of ejection energies associated with significant sources of chemical enrichment, including AGB winds ($10^{46}$~erg)\footnote{Assuming full thermalization of the total mechanical energy output of an AGB wind.}, NS-NS mergers ($10^{49} - 10^{50}$~erg), supernovae (10$^{51}$~erg), and exotic enrichment sources, such as hypernovae, that can reach much higher energies (10$^{52}$~erg). Each run contains only sources from a single event type, as indicated in the run-name by the log of the injection energy in ergs. For example, the run beginning at 180~Myr with AGB-like events is labeled ``\runone E46". The metal enrichment from each source is tracked and evolved with a unique passive scalar tied to the individual source and separate from any additional chemical enrichment that may occur within the galaxy over time.

@@

\begin{figure*}
\centering
\includegraphics[width=0.45\linewidth]{figures/ch4/AGB1_enrichment_evolution_average_CGM}
\includegraphics[width=0.45\linewidth]{figures/ch4/NSNS1_enrichment_evolution_average_CGM} \\
\includegraphics[width=0.45\linewidth]{figures/ch4/SNE1_enrichment_evolution_average_CGM}
\includegraphics[width=0.45\linewidth]{figures/ch4/HNE1_enrichment_evolution_average_CGM}
\caption{Time evolution of the fraction of metals in each phase of the ISM (colored lines), the galaxy's disk (black, solid), and the CGM (black, dashed) as averaged across all events at a given $E_{\rm ej}$. The fractions are all normalized to the total amount of metals initially injected in each event, and the individual ISM phases (colored lines) sum to the black, solid line.}
\label{ch4:fig:ISM_CGM}
\end{figure*}

We summarize the source-averaged results for all enrichment events in runs \runone E46, \runone E49, \runone E50, \runone E51, and \runone E52 in Figure~\ref{ch4:fig:ISM_CGM} by showing the fraction of source metals contained within each phase of the ISM (colored lines, which sum to the black dashed line), and the CGM (black, solid). These plots show a clear, immediate trend across $E_{\rm ej}$ for all lines in the figure, but the differences are most striking for the ISM (black, solid), CGM (black, dashed) lines. In Chapter~\ref{ch:chapter3}
% \citep{Emerick2019}
we demonstrated that there was a significant difference in metal ejection fraction, $f_{\rm ej}$, for metals from AGB sources as compared to metals released in SNe. Here, we confirm that this is driven predominantly by differences in the energy of the events, and not necessarily when or where they occur in the ISM. The clear trend in Figure~\ref{ch4:fig:ISM_CGM} shows that events with higher $E_{\rm ej}$ are much more readily ejected from the disk of the galaxy, while events with lower $E_{\rm ej}$ couple poorly to the same galactic outflows.
%\textit{statement on how immediate enrichment is after events}
% By 10~Myr after the enrichment events,
The high-energy events rapidly converge on a peak $f_{\rm ej}$ within about 20~Myr of the event, with only a gradual increase towards the end of the 150~Myr as the metals in the ISM are swept up in additional outflows. The lower energy events evolve more gradually. The lowest energy event, corresponding to AGB winds, reaches $f_{\rm ej} \sim$~0.68, which increases to 0.87 for the 10$^{51}$~erg events, and 0.95 for the 10$^{52}$~erg event. To emphasize these differences, we plot each line separately in the left panel of Figure~\ref{ch4:fig:CGM_CNM}

This implies that, if one to were assume that all metals behaved the same and adopted an $f_{\rm ej}$ of $\sim$0.9 -- typical for the total metal ejection fraction in simulations that do not track individual enrichment channels \citep[e.g.][]{Muratov2017,Christensen2018}, and consistent with observations of the O abundance in Local Group dwarf galaxies \citep[e.g.][]{Kirby2011,McQuinn2015} -- one would underestimate the amount of AGB metals in the ISM by a factor of $\sim$ 3, and overestimate exotic, high-energy sources by a factor of $\sim$ 2. In each case, these are metals that cannot participate in the enrichment of future stellar populations and will make a significant difference on the output stellar abundances in chemical evolution models. We explore this impact in different works.

In all cases, the metals contained with the ISM are deposited predominately in the ionized phases of the ISM, the WIM or the HIM, with the relative fraction in each phase driven by the energy of the event. Metals in the ISM injected with $E_{\rm ej} > 10^{49}$~erg are initially located predominately in the HIM, roughly 1/2 for $E_{\rm ej} = 10^{49}$~erg and nearly all for $E_{\rm ej} > 10^{51}$~erg. The lowest energy events are initially in the WIM and WNM, tracing the two dominant volume-filling components of the ISM (see Chapter~\ref{ch:chapter1}), as these events do not have sufficient energy to generate the HIM by themselves.

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.45\linewidth]{figures/ch4/CGM_average_evolution}
  \includegraphics[width=0.45\linewidth]{figures/ch4/CNM_average_evolution}
  \caption{Time evolution of the fraction of enrichment source metals contained in the CGM over time (left), and the fraction of metals in the CNM just those metals retained in the ISM for each enrichment source (right). These are the same as the lines in Figure~\ref{ch4:fig:ISM_CGM}, but the CNM lines are normalized by the black, dashed total ISM line in Figure~\ref{ch4:fig:ISM_CGM}.}
  \label{ch4:fig:CGM_CNM}
\end{figure*}
% make a figure here showing JUST the CNM evolution of the ISM for each source

However, the metals from each source do gradually trend towards a fraction of $\sim$0.8 by the end of the 150~Myr simulation time, corresponding to the mass fraction of the CNM at this point in the galaxy's evolution (see Figure~\ref{ch1:fig:ISM_evolution}). This trend -- that the fraction of metals contained in a given phase tends towards the mass fraction of that phase -- is true across all phases in the simulation. Therefore, we can consider that the metals for each source are well-mixed across the \textit{phases} of the ISM on timescales of $\sim$100-150~Myr. This is comparable, but less than, the dynamical timescale of the galaxy, $\sim$200~Myr. These two results suggest that the delay-time between when metals are available to enrich ongoing star formation is on order of $\sim$10~Myr across sources, and only significant for the lowest (E46, AGB-like) energy sources. Although well-mixed across phases, we emphasize that this does not imply that the abundances are the same across phases nor that the metals are spatially well-mixed across the galaxy. We investigate this second point further below.

@@

To frame this analysis, lets first examine what processes affect the metal mass fraction PDF. The mean metal mass fraction for a collection of gas is simply the total metal mass divided by the total gas mass. For a collection of gas elements, the mean metal mass fraction is defined as $Z_{\rm gas} = \frac{ \Sigma Z_{\rm i}M_{\rm gas,i}}{\Sigma M_{\rm gas,i}}$, where $M_{\rm gas,i}$ is the mass of the i$^{\rm th}$ homogeneous blob of gas, and $Z_{i}$ its metal mass fraction. A completely homogeneous ISM would have a $\delta$ function metal mass fraction PDF at the mean metal fraction, or $p(Z) = \delta(Z_{\rm gas})$. By definition, then, a homogeneous distribution will have a zero-width IQR, IDR, and mean-median difference. Non-uniform enrichment in a homogeneous medium will increase each of these quantities, while mixing in a non-uniform medium will tend to decrease each. In simple models of galactic chemical evolution, the mean gas abundance -- not the median -- is followed, even though the median is more indicative of the abundance of a ``typical" gas element. Because these two quantities provide a more immediate conceptual understanding of how metals are distributed in the ISM, we focus on characterizing the inhomogeneity of the followed metals by examining the mean-median difference. Though it is necessary for this quantity to be zero for metals to be homogeneously distributed in a medium, we do note that it is not a sufficient condition (the mean of a symmetric distribution is equal to the median). However, its evolution is easy to interpret conceptually. Enrichment increases this quantity by raising the mean metal fraction while keeping the median value fixed (or only slightly increasing), so long as the newly enriched mass is a small fraction of the total gas mass (as is almost always the case). Conversely, preferential removal of metals from a medium will lower this quantity. And finally, the mixing of metal-rich gas elements with metal-poor will gradually bring the mean and median values to parity. For our numerical experiments, enrichment in the whole ISM only occurs once and ejection from the ISM is easy to follow. This allows for a cleaner interpretation of the evolution than examining either the IQR or the IDR.

@@

Clearly, at no point can any of these distributions be considered well-mixed within the 150~Myr simulation time. Prior studies of passive scalar mixing in isothermal turbulent boxes find typical mixing timescales\footnote{Defined roughly as the time it takes for the r.m.s. abundance to equilibrate, not the time it takes to completely homogenize, which tends towards infinity.} on order of the eddy turnover timescale \citep[e.g.][]{SurPanScannapieco2014} ($t_{\rm ed}$). Typically, $t_{\rm ed}$ is defined as the relevant length scale (e.g. the diameter of our galaxy) divided by the r.m.s gas velocity. Although extending these results to multi-phase, global-galaxy simulations is non-trivial, we find typical $t_{\rm ed}$ in the CNM of $\sim$100~Myr and in the HIM of $\sim$10~Myr.\footnote{For comparison, the isothermal sound speed in these phases are $\sim$4~Gyr and $\sim$10~Myr respectively.} In order to properly evaluate the mixing timescale we would need to run each simulation for significantly longer, for several $t_{\rm ed}$.

@@

What this does suggest, however, is that in utilizing stars with unusual abundances at low metallicities to infer the yields of exotic enrichment sources, the spread of abundances in those stars may provide an important insight into the origin of those metals. For example, r-process enrichment from exotic enrichment events like hypernovae should manifest itself as more well-mixed than if the elements originated in lower energy NS-NS mergers. Conducting this analysis requires measured r-process abundances for a large sample of stars in an individual dwarf galaxy. This is challenging for the lowest mass, most metal poor dwarf galaxies, which typically only have a few to 10 stars with measured abundances, but could be done for more a few low mass dwarfs like Ursa Minor, and for more massive dwarfs like Carina, and Fornax \citep{Suda2017}.
% somerthing about needing cool onezone chemical models to try and model data with many free params might be nice here.

@@

For clarity we have focused on the \textit{mean} evolution of individual enrichment events at fixed injection energy in the results thus far. We turn now to discuss how much variety exists among these individual events. For each of the discussed runs (from set \runonenu), Figure~\ref{ch4:fig:CGM_CNM_variance} shows the fraction of metals ejected from the galaxy (left) and the fraction of metals in the ISM within the CNM (center) as a function of source energy at two different times, 20~Myr (blue) and 100~Myr (orange). In the right panel, we show how the spread at each of these times varies for each event. The averaged values at each energy are shown as points, and the bars denote the min/max value for each event energy.\footnote{While the standard deviation or IQR may be more informative here, rather than min / max, the sample size at each energy, with the exception of the E46 runs, is too small to reliably do so.} As shown, there is substantial variation at each energy. This is particularly striking for the E46 events, which seemingly can have any value of $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ at any point during the evolution. Finally, the right panel shows that the degree of inhomogeneous mixing for each event varies significantly as well. However, there is a more clear trend for each of the mean, min, and max values of the spread at 100~Myr (decreasing as a function of increasing $E_{\rm ej}$).

@@

It is clear that the trends discussed thus far applies to the average behavior of events at a fixed $E_{\rm ej}$, but not for an individual event. In addition, a higher $E_{\rm ej}$ seems to increase the minimum $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ expected, but the behavior of any one event must depend on additional factors. But what causes this variation among individual events? Do additional properties about the injection site (ISM density, position, local SFR), determine the subsequent evolution? Or is this the result of a non-deterministic set of complex interactions the metals from a given event experience over their lifetime in the galaxy? We investigate the dependence on a few of these properties below.

\subsection{Dependence on Radial Position}
\label{ch4:sec:radial position}
We place each source at regular, but arbitrary, positions in the galaxy without consideration for the local ISM conditions at injection. Sources are placed at the center of the galaxy ($r = 0$~pc) or at various radii (at cardinal positions at 300~pc, and 600~pc), all in the mid-place of the galaxy. Since the E46 events were likely to not self-interact, we additionally placed events at $r = 100$~pc and one event at each $r$ at about one scale height above the disk ($z = 50$~pc). For each \runonenu~run, we plot $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ as a function of radial position in the galaxy for each event in Figure~\ref{ch4:fig:radial position}. The points, error bars, and colors are the same as in Figure~\ref{ch4:fig:CGM_CNM_variance}.

@@

Across runs, there is no clear dependence on the radial position of the event and the fraction of metals ejected beyond the galaxy for any of the displayed runs at either 20~Myr or 100~Myr. There is a slight increase in the minimum fraction of metals ejected from the galaxy as a function of radius for the higher energy runs ($>10^{50}$~erg). This is perhaps not surprising for the largest radii as 600~pc is right near the very edge of the galaxy's disk. But again, the trend is weak and there is clearly still a great variance at both fixed $E_{\rm ej}$ and $r$. The fraction of metals in the ISM that are contained within the CNM (middle row) do not show any obvious trend at E46, but does seem to generally be lowest for the higher energy sources at the outskirts of the galaxy. Finally, the radial position does seem to affect the mixing behavior. Sources at the outskirts of the galaxy tend to be more inhomogeneous than sources at inner radii, especially at 20~Myr for the higher energy events, and the variation in behavior for these sources decreases substantially, particularly at 100~Myr. It is unclear what exactly drives these mixing differences. If feedback-driven turbulence is the dominant driver of ISM mixing in this dwarf galaxy -- as it likely is -- then sources near the center are more likely to be affected by more feedback events (due to a slightly higher $\Sigma_{\rm SFR}$) and mix more efficiently.

>>>>

The elemental abundances of a galaxy over time are sensitive to the nuclear physics and stellar astrophysics that determines which stars make what elements and when. The stellar abundance patterns within a given galaxy also depend on the details of how those metals are released into the ISM through various forms of stellar feedback, and the hydrodynamic interactions that ultimately mix those elements into the ISM or eject them from the galaxy in galactic winds. Stellar abundance patterns in old, metal poor systems are the result of enrichment from substantially fewer sources than younger, more metal rich stellar populations. Because of this, studying abundance patterns in these environments allows us to place constraints on the yields of individual enrichment events and the nucleosynthetic sites of each element. The lowest mass dwarf galaxies in the Local Group, UFDs, offer some of the best constraints on these processes. These environments can potentially be used to trace and constrain enrichment from the first stars \citep[e.g.][]{FrebelBromm2012,Ji2015,Ritter2015,Jeon2017,Hartwig2018} and have been used to place constraints on the variety of possible astrophysical sources of r-process enrichment \citep[e.g.][]{Ji2016a,Ji2016b,Ji2018a,Ji2019,Tsujimoto2017,Duggan2018,Nagasawa2018,Ojima2018}. It is likely that chemical enrichment in more massive, metal rich systems with ongoing star formation like the Milky Way exhibits smaller abundance spreads, averaging over the variety of complex hydrodynamics and feedback-driven turbulence that may be driving much of the significant variance in stellar abundances seen at low metallicity. Stochasticity, which has been shown to be important for setting the width of stellar abundance patterns in low metallicity environments of the Milky Way halo \citep[e.g.][]{Cescutti2008,Cescutti2014}, plays an even greater role in the evolution of these low mass dwarf galaxies \citep{Applebaum2018}. Indeed the increase in scatter in the stellar abundance patterns of low mass dwarf galaxies and UFDs has been attributed to inhomogeneous mixing and stochastic effects \citep[e.g.][]{Norris2010a,Lee2013,Simon2015,Mashonkina2017,Suda2017}. This complicates the interpretation of abundance patterns in these galaxies, particularly for constraining individual enrichment events.

@@

In this Chapter we utilize the simulations analyzed in prior chapters to conduct a controlled set of ``mixing experiments'' whereby we restart each simulation with enrichment events placed by-hand in order to more directly investigate the evolution of metals from an individual event. While elemental yields were tracked for each star in our simulation, we lacked the necessary Lagrangian information\footnote{The history of a given fluid element, which is followed natively in SPH codes, but not in the grid-based hydrodynamics simulations used here.} about the metals once they were released into the ISM to be able to trace the evolution of single enrichment events. We investigate primarily how the feedback ejection energy of individual sources ($E_{\rm ej}$) and global star formation rate at the time of enrichment affects how metals are ejected from the galaxy in galactic winds and metal mixing in the ISM. In Section~\ref{ch4:sec:methods} we briefly outline our methods and discuss the setup of these mixing experiments. In Section~\ref{ch4:sec:results} we discuss the results of these experiments and the role of $E_{\rm ej}$, global SFR, radial position of the enrichment event, and local ISM density around each enrichment sources affects the evolution of its metals, and discuss and conclude in Section~\ref{ch4:sec:discussion conclusion}.

@@

\subsection{Mixing Experiment Setup}
\label{ch4:sec:experiment}
We restart our fiducial, full-physics simulation at two different times, 180 Myr, and 360 Myr, labeled as runs \runonenu~and \runtwonu~respectively. These correspond to two different times in the galaxy's SFR evolution, testing how much variance is expected in the metal mixing and ejection with the star formation rate. Run \runonenu~occurs during the lull in star formation following the initial SFR peak, and run \runtwonu~occurs in an extended period of little to no ongoing star formation. We attempted to to evolve each simulation for 150 Myr, but due to computational constraints this was not always possible.\footnote{We began runs during the peak of star formation ($\sim 100$~Myr), but the computational cost increases substantially during this time. These runs were abandoned.}

@@

At the beginning of each restart, we place by-hand one or more enrichment events at assigned positions throughout the galaxy, with thermal injection energies ($E_{\rm ej}$), masses ($m_{\rm ej}$), and metal fractions ($Z_{\rm ej}$). We note that the injection masses, particularly the metal fractions, are somewhat arbitrary. The injected metal masses are dynamically insignificant relative to the ambient ISM mass in which they occur; these fields, therefore, act as tracers and do not impact the dynamics or cooling physics of the simulations. The important parameter here is $E_{\rm ej}$, which we vary to sample the range of ejection energies associated with significant sources of chemical enrichment, including AGB winds ($10^{46}$~erg)\footnote{Assuming full thermalization of the total mechanical energy output of an AGB wind.}, NS-NS mergers ($10^{49} - 10^{50}$~erg), supernovae (10$^{51}$~erg), and exotic enrichment sources, such as hypernovae, that can reach much higher energies (10$^{52}$~erg). Each run contains only sources from a single event type, as indicated in the run-name by the log of the injection energy in ergs. For example, the run beginning at 180~Myr with AGB-like events is labeled ``\runone E46". The metal enrichment from each source is tracked and evolved with a unique passive scalar tied to the individual source and separate from any additional chemical enrichment that may occur within the galaxy over time.

@@

\begin{figure*}
\centering
\includegraphics[width=0.45\linewidth]{figures/ch4/AGB1_enrichment_evolution_average_CGM}
\includegraphics[width=0.45\linewidth]{figures/ch4/NSNS1_enrichment_evolution_average_CGM} \\
\includegraphics[width=0.45\linewidth]{figures/ch4/SNE1_enrichment_evolution_average_CGM}
\includegraphics[width=0.45\linewidth]{figures/ch4/HNE1_enrichment_evolution_average_CGM}
\caption{Time evolution of the fraction of metals in each phase of the ISM (colored lines), the galaxy's disk (black, dashed), and the CGM (black, solid) as averaged across all events at a given $E_{\rm ej}$. The fractions are all normalized to the total amount of metals initially injected in each event, and the individual ISM phases (colored lines) sum to the black, solid line.}
\label{ch4:fig:ISM_CGM}
\end{figure*}

We summarize the source-averaged results for all enrichment events in runs \runone E46, \runone E49, \runone E50, \runone E51, and \runone E52 in Figure~\ref{ch4:fig:ISM_CGM} by showing the fraction of source metals contained within each phase of the ISM (colored lines, which sum to the black dashed line), and the CGM (black, solid). These plots show a clear, immediate trend across $E_{\rm ej}$ for all lines in the figure, but the differences are most striking for the ISM (black, dashed), CGM (black, solid) lines. In Chapter~\ref{ch:chapter3}
% \citep{Emerick2019}
we demonstrated that there was a significant difference in metal ejection fraction, $f_{\rm ej}$, for metals from AGB sources as compared to metals released in SNe. Here, we confirm that this is driven predominantly by differences in the energy of the events, and not necessarily when or where they occur in the ISM. The clear trend in Figure~\ref{ch4:fig:ISM_CGM} shows that events with higher $E_{\rm ej}$ are much more readily ejected from the disk of the galaxy, while events with lower $E_{\rm ej}$ couple poorly to the same galactic outflows.
%\textit{statement on how immediate enrichment is after events}
% By 10~Myr after the enrichment events,
The high-energy events rapidly converge on a peak $f_{\rm ej}$ within about 20~Myr of the event, with only a gradual increase towards the end of the 150~Myr as the metals in the ISM are swept up in additional outflows. The lower energy events evolve more gradually. The lowest energy event, corresponding to AGB winds, reaches $f_{\rm ej} \sim$~0.68, which increases to 0.87 for the 10$^{51}$~erg events, and 0.95 for the 10$^{52}$~erg event. To emphasize these differences, we plot each line separately in the left panel of Figure~\ref{ch4:fig:CGM_CNM}

This implies that, if one to were assume that all metals behaved the same and adopted an $f_{\rm ej}$ of $\sim$0.9 -- typical for the total metal ejection fraction in simulations that do not track individual enrichment channels \citep[e.g.][]{Muratov2017,Christensen2018}, and consistent with observations of the O abundance in Local Group dwarf galaxies \citep[e.g.][]{Kirby2011,McQuinn2015} -- one would underestimate the amount of AGB metals in the ISM by a factor of $\sim$ 3, and overestimate exotic, high-energy sources by a factor of $\sim$ 2. In each case, these are metals that cannot participate in the enrichment of future stellar populations and will make a significant difference to the output stellar abundances in chemical evolution models.
% We explore this impact in different works.

In all cases, the metals contained within the ISM are deposited predominantly in the ionized phases of the ISM, the WIM or the HIM, with the relative fraction in each phase driven by the energy of the event. Metals in the ISM injected with $E_{\rm ej} > 10^{49}$~erg are initially located predominately in the HIM, roughly 1/2 for $E_{\rm ej} = 10^{49}$~erg and nearly all for $E_{\rm ej} > 10^{51}$~erg. The lowest energy events are initially in the WIM and WNM, tracing the two dominant volume-filling components of the ISM (see Chapter~\ref{ch:chapter1}), as these events do not have sufficient energy to generate the HIM by themselves.

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.45\linewidth]{figures/ch4/CGM_average_evolution}
  \includegraphics[width=0.45\linewidth]{figures/ch4/CNM_average_evolution}
  \caption{Time evolution of the fraction of enrichment source metals contained in the CGM over time (left), and the fraction of metals in the CNM relative to just those metals retained in the ISM for each enrichment source (right). These are the same as the lines in Figure~\ref{ch4:fig:ISM_CGM}, but the CNM lines are normalized by the black, dashed total ISM line in Figure~\ref{ch4:fig:ISM_CGM}.}
  \label{ch4:fig:CGM_CNM}
\end{figure*}
% make a figure here showing JUST the CNM evolution of the ISM for each source

However, the metals from each source do gradually trend towards a fraction of $\sim$0.8 by the end of the 150~Myr simulation time. This value corresponds to the total mass fraction of the CNM (see Figure~\ref{ch1:fig:ISM_evolution}). This trend -- that the fraction of metals contained in a given phase tends towards the mass fraction of that phase -- is true across all phases in the simulation. Therefore, we can consider that the metals for each source are well-mixed across the \textit{phases} of the ISM on timescales of $\sim$100-150~Myr. This is comparable to, but less than, the dynamical timescale of the galaxy, $\sim$200~Myr. These two results suggest that the delay-time between when metals are available to enrich ongoing star formation is on order of $\sim$10~Myr across sources, and only significant for the lowest (E46, AGB-like) energy sources. Although well-mixed across phases, we emphasize that this does not imply that the abundances are the same across phases nor that the metals are spatially well-mixed across the galaxy. We investigate this second point further below.

@@

To frame this analysis, let us first examine what processes affect the metal mass fraction PDF. The mean metal mass fraction for a collection of gas is simply the total metal mass divided by the total gas mass. For a collection of gas elements, the mean metal mass fraction is defined as $Z_{\rm gas} =  \Sigma Z_{\rm i}M_{\rm gas,i} / \Sigma M_{\rm gas,i}$, where $M_{\rm gas,i}$ is the mass of the i$^{\rm th}$ homogeneous blob of gas, and $Z_{i}$ its metal mass fraction. A completely homogeneous ISM would have a $\delta$ function metal mass fraction PDF at the mean metal fraction, or $p(Z) = \delta(Z_{\rm gas})$. By definition, then, a homogeneous distribution will have a zero-width IQR, IDR, and mean-median difference. Non-uniform enrichment in a homogeneous medium will increase each of these quantities, while mixing in a non-uniform medium will tend to decrease each. In simple models of galactic chemical evolution, the mean gas abundance -- not the median -- is followed, even though the median is more indicative of the abundance of a ``typical" gas element. Because these two quantities provide a more immediate conceptual understanding of how metals are distributed in the ISM, we focus on characterizing the inhomogeneity of the followed metals by examining the mean-median difference. Though it is necessary for this quantity to be zero for metals to be homogeneously distributed in a medium, we do note that it is not a sufficient condition (the mean of a symmetric distribution is equal to the median). However, its evolution is easy to interpret conceptually. Enrichment increases this quantity by raising the mean metal fraction while keeping the median value fixed (or only slightly increasing), so long as the newly enriched mass is a small fraction of the total gas mass (as is almost always the case). Conversely, preferential removal of metals from a medium will lower this quantity. And finally, the mixing of metal-rich gas elements with metal-poor will gradually bring the mean and median values to parity. For our numerical experiments, enrichment in the whole ISM only occurs once and ejection from the ISM is easy to follow. This allows for a cleaner interpretation of the evolution than examining either the IQR or the IDR.

@@

Clearly, at no point can any of these distributions be considered well-mixed within the 150~Myr simulation time. Prior studies of passive scalar mixing in isothermal turbulent boxes find typical mixing timescales\footnote{Defined roughly as the time it takes for the r.m.s. abundance to equilibrate, not the time it takes to completely homogenize, which tends towards infinity.} on order of the eddy turnover timescale \citep[e.g.][]{SurPanScannapieco2014} ($t_{\rm ed}$). Typically, $t_{\rm ed}$ is defined as the relevant length scale (e.g. the diameter of our galaxy) divided by the r.m.s gas velocity. Although extending these results to multi-phase, global-galaxy simulations is non-trivial, we find typical $t_{\rm ed}$ in the CNM of $\sim$100~Myr and in the HIM of $\sim$10~Myr.\footnote{For comparison, the isothermal sound crossing time in these phases are $\sim$4~Gyr and $\sim$10~Myr respectively.} In order to properly evaluate the mixing timescale we would need to run each simulation for significantly longer, for several $t_{\rm ed}$.

@@

What this does suggest, however, is that in utilizing stars with unusual abundances at low metallicities to infer the yields of exotic enrichment sources, the spread of abundances in those stars may provide an important insight into the origin of those metals. For example, r-process enrichment from exotic enrichment events like hypernovae should manifest itself as more well-mixed than if the elements originated in lower energy NS-NS mergers. Conducting this analysis requires measured r-process abundances for a large sample of stars in an individual dwarf galaxy. This is challenging for the lowest mass, most metal poor dwarf galaxies, which typically only have a few to 10 stars with measured abundances, but could be done for low mass dwarfs like Ursa Minor, and for more massive dwarfs like Carina, and Fornax \citep{Suda2017}.
% somerthing about needing cool onezone chemical models to try and model data with many free params might be nice here.

@@

For clarity we have focused on the \textit{mean} evolution of individual enrichment events at fixed injection energy in the results thus far. We turn now to discuss how much variety exists among these individual events. For each of the discussed runs (from set \runonenu), Figure~\ref{ch4:fig:CGM_CNM_variance} shows the fraction of event tracer metals ejected from the galaxy (left) and the fraction of metals in the ISM within the CNM (center) as a function of source energy at two different times, 20~Myr (blue) and 100~Myr (orange). In the right panel, we show how the spread at each of these times varies for each event. The averaged values at each energy are shown as points, and the bars denote the min/max value for each event energy.\footnote{While the standard deviation or IQR may be more informative here, rather than min / max, the sample size at each energy, with the exception of the E46 runs, is too small to reliably do so.} As shown, there is substantial variation at each energy. This is particularly striking for the E46 events, which seemingly can have any value of $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ at any point during the evolution. Finally, the right panel shows that the degree of inhomogeneous mixing for each event varies significantly as well. However, there is a more clear trend for each of the mean, min, and max values of the spread at 100~Myr (decreasing as a function of increasing $E_{\rm ej}$).

@@

It is clear that the trends discussed thus far apply to the average behavior of events at a fixed $E_{\rm ej}$, but not for an individual event. In addition, a higher $E_{\rm ej}$ seems to increase the minimum $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ expected, but the behavior of any one event must depend on additional factors. But what causes this variation among individual events? Do additional properties about the injection site (ISM density, position, local SFR), determine the subsequent evolution? Or is this the result of a non-deterministic set of complex interactions the metals from a given event experience over their lifetime in the galaxy? We investigate the dependence on a few of these properties below.

\subsection{Dependence on Radial Position}
\label{ch4:sec:radial position}
We place each source at regular, but arbitrary, positions in the galaxy without consideration for the local ISM conditions at injection. Sources are placed at the center of the galaxy ($r = 0$~pc) or at various radii (at cardinal positions at 300~pc, and 600~pc), all in the mid-plane of the galaxy. Since the E46 events were likely to not self-interact, we additionally placed events at $r = 100$~pc and one event at each $r$ at about one scale height above the disk ($z = 50$~pc). For each \runonenu~run, we plot $f_{\rm CGM}$ and $f_{\rm CNM} / f_{\rm Disk}$ as a function of radial position in the galaxy for each event in Figure~\ref{ch4:fig:radial position}. The points, error bars, and colors are the same as in Figure~\ref{ch4:fig:CGM_CNM_variance}.

@@

Across runs, there is no clear dependence on the radial position of the event and the fraction of metals ejected beyond the galaxy for any of the displayed runs at either 20~Myr or 100~Myr. There is a slight increase in the minimum fraction of metals ejected from the galaxy as a function of radius for the higher energy runs ($>10^{50}$~erg). This is perhaps not surprising for the largest radii as 600~pc is right near the very edge of the galaxy's disk. But again, the trend is weak and there is clearly still a great variance at both fixed $E_{\rm ej}$ and $r$. The fraction of metals in the ISM that are contained within the CNM (middle row) does not show any obvious trend at E46, but does seem to generally be lowest for the higher energy sources at the outskirts of the galaxy. Finally, the radial position does seem to affect the mixing behavior. Sources at the outskirts of the galaxy tend to be more inhomogeneous than sources at inner radii, especially at 20~Myr for the higher energy events, and the variation in behavior for these sources decreases substantially, particularly at 100~Myr. It is unclear what exactly drives these mixing differences. If feedback-driven turbulence is the dominant driver of ISM mixing in this dwarf galaxy -- as it likely is -- then sources near the center are more likely to be affected by more feedback events (due to a slightly higher $\Sigma_{\rm SFR}$) and mix more efficiently.
